Structured Synchronous Reactive Programming for Game Development \\ \Large{Case Study: On Rewriting Pingus from C++ to \CEU}

### 2017-05-24 paper.tex
...

In Pingus, the game logic also accounts for almost half the size of the
codebase: 18.173 from 39.362 LoCs (46\%) spread across 272 files.
%
However, about half of the game logic relates to non-reactive code, such as
configurations and options, saved games and serialization, maps and levels
descriptions, string formatting, collision detection, graph algorithms, etc.
This part remains unchanged and relies on the seamless integration between \CEU
and C/C++~\cite{ceu.sensys13}.
%
Therefore, we rewrote 9.186 LoCs spread across 126 files%
\footnote{Complete codebase: \url{github.com/an000/p/tree/master/cpp}}.
%
In order to only consider effective code in the analysis, we then removed all
headers, declarations, trivial getters \& setters, and other innocuous
statements, resulting 4.135 dense LoCs spread across 70 implementation files
originally written in C++%
\footnote{C++ codebase: \url{github.com/an000/p/tree/master/all}}.
We did the same with the implementation in \CEU, resulting in 3.697 dense LoCs%
\footnote{\CEU codebase: \url{github.com/an000/p/tree/master/all}}.
%
Figure~\ref{tab.tree} summarizes the effective codebase in the two
implementations.

@@

@@

Functional reactive programming is an
FRP
    - interfacing with C
    - live porting

different
    - generally?
        - mostram exemplos abstratos
            - imagine um jogo...
                - assim poderiamos aplicar...
            - sem codigo real
            - nao aplicam a jogos existentes
    - propose solutions
        - eliminate
visitor, adopted in the language

- patterns
    - sbgames
    - gpp   
- contol-flow patterns that are recurrent in Pingus and likely apply to other
games
- also, not identifying and applying, but applying in idiomatic code in another
  language

    - como eh diferente de GoF
        - not much unlike behavioral patterns

Design patterns for games can be approached from two dif-
ferent perspectives; firstly as patterns used for describing
the game mechanics (gameplay and game rules) and sec-
ondly as the use of object-oriented design patterns in pro-
gramming games.
Concerning game mechanics, Bjork et al. [4] introduced
a set of design patterns which essentially are descriptions
(employing a unified vocabulary) of reoccurring interac-
tion schemes relevant to game’s story and gameplay. As
such, these patterns are not related to the software archi-
tecture or code. The proposed patterns are collected from
interviewing professional game programmers, from ana-
lyzing existing games and from transforming game
mechanics.

TODO:
    - FSMs
    - BTs
    - FRP

>>>>

In Pingus, the game logic also accounts for almost half the size of the
codebase: 18.173 from 39.362 LoCs (46\%) spread across 272 files.
%
However, about half of the game logic relates to non-reactive code, such as
configurations and options, saved games and serialization, maps and levels
descriptions, string formatting, collision detection, graph algorithms, etc.
This part remains unchanged and relies on the integration between \CEU and
C/C++.
%
Therefore, we rewrote 9.186 LoCs spread across 126 files%
\footnote{Complete codebase: \url{github.com/an000/p/tree/master/cpp}}.
%
In order to only consider effective code in the analysis, we then removed all
headers, declarations, trivial getters \& setters, and other innocuous
statements, resulting 4.135 dense LoCs spread across 70 implementation files
originally written in C++%
\footnote{C++ codebase: \url{github.com/an000/p/tree/master/all}}.
We did the same with the implementation in \CEU, resulting in 3.697 dense LoCs%
\footnote{\CEU codebase: \url{github.com/an000/p/tree/master/all}}.
%
Figure~\ref{tab.tree} summarizes the effective codebase in the two
implementations.

@@

We employed \emph{live code translation}, i.e., starting from the original
codebase in C++, we reimplemented piece-by-piece without breaking the game
compilation and execution.
%
This was only possible given the seamless integration between \CEU and
C/C++~\cite{ceu.sensys13}: the type systems are equivalent and the integration
happens at the source code level.
This enables trivial sharing of control (calling C/C++ from \CEU and
vice-versa) and data (accessing C/C++ data from \CEU and vice-versa).

@@

Functional reactive programming (FRP)~\cite{frp.fran} contrasts with
Structured synchronous reactive programming as a complementary programming
style for reactive applications.
%
We believe that FRP is more suitable for data-intensive applications, while 
SSRP, for control-intensive applications.
%
On the one hand, FRP uses declarative formulas to specify continuous functions 
over time, such as for physics or data constraints among entities.
%
On the other hand, describing a sequence of steps in FRP requires to encode 
explicit state machines so that functions can switch behavior depending on the 
current state.
%
FRP has been successfully used to implement a 3D first person shooting game
from scratch, but with performance considerations~\cite{games.frag}.
%
We rewrote an existing game in small steps while keeping it working.
Although we do not provide a performance evaluation (Pingus is not performance
sensitive), previous work on \CEU shows that it is comparable to C in the
context of embedded systems~\cite{ceu.sensys13}.
Nonetheless, given the tight integration between with C/C++, critical parts of
games can be preserved in C++ if needed.
